\documentclass[../../book.tex]{subfiles}

\begin{document}
The goal of this section is to introduce \emph{solvability} of groups,
which plays an essential role in the insolvability of general quintics. 
% Def - Normal Series, Factor Groups, Length of Series
% Def - Refinement, Proper Refinement, Equivalence, Composition Series.
\begin{dfn} Normal Series, Factor Groups, Length 
and (Proper) Refinements of a Series, Equivalence of Series, Composition Series.
    
    Let $G$ be a group and 
    $S = (S_i)_{i\leq n}$ a finite sequence of subgroups.
    Then $S$ is called a \textbf{normal series in $G$} when the following conditions are met.
    \begin{enumerate}
        \item $S_0$ is the trivial subgroup and $S_n$ is $G$. 
        \item For all $i < n$, $S_i \normsub S_{i+1}$.
    \end{enumerate}
    For a normal series $S$, 
    its \textbf{factor groups}, or simply \textbf{factors},
    are groups in the sequence of quotients $(S_{i+1}/S_i)_{i<n}$.
    
    The \textbf{length} of a normal series $S$ is
    the number of strict inclusions in $S$, which we shall denote with $|S|$.
    Equivalently, it's the number of non-trivial factor groups from $S$. 
    
    Let $S$ and $T$ be normal series of $G$. 
    Then $S \leq T$ is defined as $S$ being a subsequence of $T$. 
    We say $T$ is a \textbf{refinement of $S$}.
    It is easy to see that if $S \leq T$, then $|S| \leq |T|$.
    We say $T$ is a \textbf{proper refinement of $S$}
    when it is a refinement with strictly longer length. 
    
    For a normal series $S$, 
    the following are equivalent by the 3rd isomorphism theorem.
    \begin{enumerate}
        \item $S$ has no proper refinements.
        \item For a group, it is called \textbf{simple} when
        the only normal subgroups are itself and the trivial subgroup.
        Then the factor groups of $S$ are all simple. 
    \end{enumerate}
    If either of the above is true, 
    $S$ is called a \textbf{composition series} of $G$. 
    
    For two normal series $S, T$ of $G$, 
    they are called \textbf{equivalent} when 
    there exists a bijection between the \emph{non-trivial} factors of $S$ and $T$
    that maps factors to isomorphic factors. 
    This unsurprisingly forms an equivalence relation 
    on the set of all normal series of a group $G$. 
    Note that if $T$ is a non-proper refinement of $S$, 
    then $T$ is equivalent to $S$. 
    
\end{dfn}
\begin{thm} Existence of Composition Series for Finite Groups.
    
    Let $G$ be a finite group.
    Then it has a composition series.
    
\end{thm}
\begin{proof}
    
    Let $S$ be a normal series of $G$.
    Clearly the length of $S$ cannot be larger than the cardinality of $G$
    and $\<e\> \normsub G$ is a normal series of $G$.
    Hence the set of lengths of normal series of $G$ is a non-empty
    bounded subset of the naturals
    and thus must have a maximum. 
    Let $\hat{S}$ be a normal series of $G$ with maximal length. 
    It cannot have a proper refinement, so it must be a composition series.
    
\end{proof}

\begin{rmk}
    
    The motivation for considering normal series up to equivalence is that
    the data of a normal series is really in its non-trivial factor groups.
    As we will now see, all finite groups have \emph{unique} composition series.
    But first, the theorem that you might have been wondering about.
    
\end{rmk}
% Thm - 2nd Iso.
\begin{thm} 2nd Isomorphism.
    
    Let $A, B$ be subgroups of a group $G$ such that
    $B$ is in the normaliser of $A$. 
    Define \[
        AB := \{ ab \mid a \in A, b \in B\}
    \]
    Then $AB$ is a subgroup of $G$ with $AB = BA$
    and we have the following isomorphism of groups. 
    \[
            \frac{AB}{A} \iso \frac{B}{A \cap B}
    \]
    
\end{thm}
\begin{proof}
    
    $AB$ being a subgroup and 
    the normality of $A \normsub AB$. $A \cap B \normsub B$
    is left as an exercise to the reader.
    
    Let $a \in A$ and $b \in B$. 
    Then $ab = b (b^{-1}ab)$ which is in $BA$ since $A$ is normal in $B$. 
    Also, $ba = (bab^{-1}) b$ is in $AB$ for the same reason.
    Thus $AB = BA$. 
    
    As for the isomorphism, it is easily deduced from the 1st isomorphism theorem.
    
\end{proof}
% Lem - Butterfly / Zassenhaus.
\begin{lem} Butterfly / Zassenhaus.
    
    Let $G$ be a group with subgroups $A, A^*, B, B^*$ such that 
    $A^* \normsub A$ and $B^* \normsub B$. 
    Then \begin{itemize}
        \item $A^* (A \cap B^*) \normsub A^* (A \cap B)$
        \item $(A^* \cap B) B^* \normsub (A \cap B) B^*$
        \item \[
            \frac {A^* (A \cap B)}{A^* (A \cap B^*)}
            \iso \frac{A \cap B}{(A^* \cap B)(A \cap B^*)}
            \iso \frac{(A \cap B)B^*}{(A^*\cap B)B^*}
        \]
    \end{itemize}
\end{lem}
\begin{proof}
    
    Since the second isomorphism is entirely analogous to the first,
    but we have two proofs of the first,
    we give different proofs for the two isomorphisms.
    Note that $A^* \normsub A$ and $B^* \normsub B$
    implies $A^* \cap B \normsub A \cap B$ and $A \cap B^* \normsub A \cap B$
    which in turn implies $(A^* \cap B)(A \cap B^*)$ is a subgroup
    by the 2nd isomorphism theorem,
    and it is normal in $A \cap B$. 
    
    The first:
    
    We first show that $A^* (A \cap B^*) \normsub A^* (A \cap B)$. 
    Let $a, \al \in A^*$, $b \in A \cap B$ and $\be \in A \cap B^*$. 
    Then \[
        ab \al\be (ab)^{-1} = ab \al\be b^{-1}a^{-1}
        = a (b\al b^{-1}) 
        [ (b\be b^{-1}) a^{-1} (b\be b^{-1})^{-1} ]
        (b\be b^{-1})
    \]
    where $a$, $b\al b^{-1}$ and the square bracket term are in $A^*$
    because $b\be b^{-1}$ is in $A \cap B^*$.
    Hence for all $x \in A^* (A \cap B)$, 
    $x (A^* (A \cap B^*)) x^{-1} \subseteq A^* (A \cap B^*)$,
    i.e. we have the normality. 
    
    We then invite the reader to stare intensely
    until they realise the following are true \[
        A^* (A \cap B) = A^* (A \cap B^*) (A \cap B) \,\,\,\,\,\,\,\,\,\,
        A \cap B \cap A^* (A \cap B^*) = (A^* \cap B)(A \cap B^*)
    \]
    and thus deduced from the the second isomorphism theorem, \[
        \frac {A^* (A \cap B)}{A^* (A \cap B^*)}
        = \frac {A^* (A \cap B^*)(A \cap B)}{A^* (A \cap B^*)}
        \iso \frac{A \cap B}{A \cap B \cap A^*(A \cap B^*) }
        = \frac{A \cap B}{(A^* \cap B)(A \cap B^*)}
    \]
    
    The second:
    
    We construct an explicit group morphism 
    from $(A \cap B)B^*$ to the quotient $(A \cap B) / (A^* \cap B)(A \cap B^*)$
    that is surjective and has $(A^* \cap B)B^*$ as the kernel,
    thus concluding the result from the 1st isomorphism theorem. 
    
    The group morphism is \[
        \phi : (A \cap B)B^* \to \frac{A \cap B}{(A^* \cap B)(A \cap B^*)},
        a b \mapsto a (A^* \cap B)(A \cap B^*)
    \]
    We must show this is well defined.
    Let $a, \al \in A \cap B$ and $b, \be \in B^*$ such that $ab = \al \be$. 
    Then $\al = a b \be^{-1}$ where $b \be^{-1} = a^{-1} \al \in A \cap B^*$
    which is a subset of $(A^* \cap B)(A \cap B^*)$.
    Hence $\al$ and $a$ are in the same coset of $(A^* \cap B)(A \cap B^*)$,
    i.e. the group morphism is well-defined as a function. 
    Furthmore, $a b \al \be = a \al (\al^{-1} b \al) \be$ 
    where $\al^{-1} b \al$ is in $B^*$ since $B^*$ is normal in $B$.
    Thus $a b \al \be = a \al (\al^{-1} b \al) \be$ is mapped to 
    $(a \al)(A^* \cap B)(A \cap B^*)$,
    proving the function to be a group morphism. 
    
    We leave surjectivity and $\ker \phi = (A^* \cap B) B^*$ 
    as an easy exercise for the reader to check. 
    
\end{proof}
% Thm - Schreier Refinement. 
\begin{thm} Schreier Refinement. 
    
    Let $S, T$ be normal series of a group $G$. 
    Then there exists normal series $\hat{S}, \hat{T}$ such that
    $S \leq \hat{S}$, $T \leq \hat{T}$ and $\hat{S}, \hat{T}$ equivalent. 
    
\end{thm}
\begin{proof}
    
    Let $S = (S_i)_{i \leq n}$ and $T = (T_j)_{j \leq m}$. 
    For $i < n, j \leq m$, define \[
        S_{i,j} := S_i (S_{i+1} \cap T_j)
    \]
    and for $i \leq n, j < m$, define \[
        T_{i,j} := (S_{i} \cap T_{j+1}) T_j
    \]
    Then by the Butterfly lemma, 
    $S_{i,j} \normsub S_{i,j+1}$ and $T_{i,j} \normsub T_{i+1,j}$
    with \[
        \frac{S_{i,j+1}}{S_{i,j}} \iso \frac{T_{i+1,j}}{T_{i,j}}
    \]
    Thus we have the following two normal series \begin{align*}
        \<e\> = S_{0,0} \normsub \cdots \normsub S_{0,m} 
        = S_{1,0} \normsub \cdots \normsub S_{1,m}
        = S_{2,0} \normsub \cdots \normsub S_{n-2,m}
        = S_{n-1,0} \normsub \cdots \normsub S_{n-1,m} = G
        \\
        \<e\> = T_{0,0} \normsub \cdots \normsub T_{n,0} 
        = T_{0,1} \normsub \cdots \normsub T_{n,1} 
        = T_{0,2}\normsub \cdots \normsub T_{n,m-2}
        = T_{0,m-1} \normsub \cdots \normsub T_{n,m-1} = G
    \end{align*}
    which are clearly refinements of $S, T$ respectively.
    These are equivalent since via the Butterfly lemma, 
    the non-trivial factors are mapped to isomorphic non-trivial factors bijectively.
    
\end{proof}
\begin{eg} 
    
    Let $G = C_6 = \<a\>$, a cyclic group generated by $a$ with order $6$. 
    Consider the following two normal series: \begin{align*}
        \<e\> \normsub \<a^2\> \normsub \<a\> \\
        \<e\> \normsub \<a^3\> \normsub \<a\>
    \end{align*}
    Note that these are in fact two composition series. 
    Then following the proof of the Schreier refinement theorem, 
    we obtain the two refinements, \begin{figure}[H]
        \centering
        \begin{tikzcd}[sep = huge]
        \<e\> = \<e\>(\<a^2\>\cap\<e\>) \ar[d,"C_1"{swap}]&
        (\<e\>\cap\<a^3\>)\<e\> = \<e\> \ar[d,"C_1"]\\
        \<e\> = \<e\>(\<a^2\>\cap\<a^3\>) \ar[d,"C_3"{swap,blue}] &
        (\<a^2\>\cap\<a^3\>)\<e\> = \<e\> \ar[d,"C_2"{red}] \\
        \<a^2\> = \<e\>(\<a^2\>\cap\<a\>) \ar[d,"="{swap}] &
        (\<a\>\cap\<a^3\>)\<e\> = \<a^3\> \ar[d,"="] \\
        \<a^2\>(\<a\>\cap\<e\>) \ar[d,"C_2"{swap,red}] & 
        (\<e\>\cap\<a\>)\<a^3\> \ar[d,"C_3"{blue}] \\
        \<a\> = \<a^2\>(\<a\>\cap\<a^3\>) \ar[d,"C_1"{swap,green}] &
        (\<a^2\>\cap\<a\>)\<a^3\> = \<a\> \ar[d,"C_1"{green}] \\
        \<a\> = \<a^2\>(\<a\>\cap\<a\>) &
        (\<a^2\>\cap\<a\>)\<a^3\> = \<a\> 
        \end{tikzcd}
    \end{figure}
    The arrows represent inclusions and are labelled with factor groups.
    The colouring of pairs shows the bijection of isomorphic factor groups
    given by the Butterfly lemma, as in the proof of Schreier refinement.
\end{eg}
% Thm - Uniqueness and Every Normal Series has Composition Refinement.
\begin{thm} Uniqueness of Composition Series.
    
    Let $G$ be a group with a composition series $S$.
    Then any other composition series $T$ is equivalent to $S$.
    Furthermore, any normal series $R$ has a refinement that is a composition series.
    
\end{thm}
\begin{proof}
    
    Easily deduced from the Schreier Refinement theorem.
    
\end{proof}

We arrive at our goal.

% Def - Solvability for Finite Groups, 3 Eqv.
\begin{dfn} Solvability of a Finite Group.
    
    Let $G$ be a finite group.
    Then the following are equivalent: \begin{enumerate}
        \item There exists a normal series of $G$ with abelian factor groups.
        \item There exists a normal series of $G$ 
        with factor groups of prime cardinality or one.
        \item There exists a normal series of $G$ with cyclic factor groups.
    \end{enumerate}
    It is also immediate from the first definition that the composition series
    of $G$ has abelian factor groups.
    If any of the above is true, we say $G$ is \textbf{solvable}. 
    
\end{dfn}
\begin{proof}
    
    ($1 \imp 2$)
        
        Let $S$ be a normal series of $G$ with abelian factor groups.
        Since $G$ is finite, it has a composition series,
        so there exists a refinement $\hat{S}$ of $S$ that is a composition series.
        It is left as an exercise to the reader 
        to deduce from the 3rd isomorphism theorem
        that the factor groups of $\hat{S}$ are also abelian.
        Let $H$ be a factor group of $\hat{S}$. 
        Then it is an abelian simple group. 
        If it is trivial, we are done, so let it be non-trivial.
        Then there is a prime $p$ that divides $|H|$ 
        (by the fundamental theorem of arithmetic).
        So by Cauchy's theorem, there exists a subgroup of cardinality $p$.
        $H$ abelian implies this subgroup is normal, 
        so $H$ simple implies this subgroup is either trivial or the entire group.
        It cannot be trivial since $1$ is not a prime.
        Thus $H$ is equal to the subgroup, i.e. it has prime cardinality. 
        
    ($2 \imp 3$) Easily deduced from Lagrange's theorem.
    
    ($3 \imp 1$) Trivial.
\end{proof}
A useful result.
% Lem - Solvable iff Exists Solvable Factor.
\begin{lem} Solvable if and only if Exists Solvable Factor.
    
    Let $G$ be a group.
    If $G$ is solvable then for all normal subgroups $N$,
    $N$ and $G / N$ are solvable.
    On the other hand if there exists a normal subgroup $N$ such that 
    $N$ and $G / N$ are solvable,
    then $G$ is solvable.
    
\end{lem}
\begin{proof}
    
    Left as an exercise. 
    [Hint : The Schreier Refinement theorem
    and the 3rd isomorphism theorem are your friends.]
    
\end{proof}

\end{document}