\documentclass[../book.tex]{subfiles}

\begin{document}
%    -  Def - Galois Ext, Galois Group
\begin{dfn} Galois Extension and Galois Group.
    Let $\iota_L : K \to L$ be a finite field extension. 
    $(L,\iota_L)$ is called a \textbf{Galois extension} when 
    it is normal and separable extension.
    Then the \textbf{Galois group of $(L,\iota_L)$} is defined as 
    the group of $K$-extension automorphisms of $(L,\iota_L)$, denoted
    \[ \gal{K}{(L,\iota_L)} := \aut{K}{(L,\iota_L)} \]
    If $\iota_L$ is clear, we simply write $\gal{K}{L}$. 
\end{dfn}
%    -  Thm - Galois Correspondance Part -1
\begin{thm} Galois Extension Degree equals Order of Galois Group.
    
    Let $\iota_L : K \to L$ be a Galois extension. 
    Then $[L : K] = \gal{K}{L}$
\end{thm}
\begin{proof}
    Since $L$ itself is an extension extending $\iota_L : K \to L$ that is normal,
    we have \[
        [L : K] = [L : K]_S = |\mor{K}{L}{L}| = |\aut{K}{L}| = |\gal{K}{L}|
    \]
\end{proof}
%    -  Lem - Important Lemma / Reverse Galois / Fixed Subfields
\begin{lem} The All Important Lemma.
    Let $L$ be a field, and $G$ be a finite subgroup of $\aut{\Ring}{L}$. 
    Since $G \leq_\Grp \aut{\Ring}{L} \subseteq \aut{\Set}{L}$, 
    we have a natural $G$ action on $L$ via $l \mapsto g(l)$ for $l \in L$. 
    Define \textbf{the subfield of $K$ fixed by $G$} as
    \[ L^G := \{l \in L \mid \forall g \in G, g(l)=l\} \]
    Then $\subseteq : L^G \to L$ is normal and separable with $[L:L^G] = |G|$,
    i.e. it is a Galois extension. 
    Furthermore, $G = \gal{L^G}{L}$.
\end{lem}
\begin{proof}
    We leave the reader to check that $L^G$ indeed forms a subfield. 

    Let $a \in L$. We claim that \[
        \min(a,L^G) = \prod_{b \in Orb(a)} (X - b)
    \]
    Hence $\min(a,L^G)$ splits and has no repeated roots. 
    
    We first show that the product is actually in $L^G[X]$. 
    Let $g \in G$. Then \[
        \bar{g} \prod_{b \in Orb(a)} (X - b) 
        = \prod_{b \in Orb(a)} (X - g(b)) = \prod_{b' \in Orb(a)} (X - b')
    \]
    Hence the product is fixed by all $g \in G$. 
    Writing the product as $\sum_{k\in n} \la_k X^k$ where $\la_k \in L$,
    we have $\la_k \in L^G$, i.e. the product is indeed in $L^G[X]$. 
    
    We now prove the product is irreducible in $L^G[X]$.
    The product is clearly non-zero and non-unit. 
    Let $f, g$ be polynomials over $L^G$ such that \[
        fg = \prod_{b \in Orb(a)} (X - b)
    \]
    Evaluating both sides at $a$, we have $0 = ev_a f$ or $0 = ev_a g$. 
    WLOG $ev_a f = 0$. 
    Then since $f \in L^G[X]$ implies for all $g \in G$, $\bar{g}f = f$, 
    we have \[
        0 = g(0) = g(ev_a f) = ev_{g(a)}(\bar{g} f) = ev_{g(a)} f
    \]
    Hence for all $b \in Orb(a)$, $b$ is a root of $f$, i.e.
    there exists a polynomial $h$ over $L$ such that \[
        h \prod_{b \in Orb(a)} (X - b) = f
    \]
    The by $L[X]$ is an integral domain, $1 = gh$,
    which implies $\deg g = 0$, giving $g$ is a unit in $L^G[X]$. 
    This proves irreducibility of the product in $L^G[X]$. 
    Hence it is equal to $\min(a,L^G)$. 
    
    Note that since we defined normal and separable extensions as finite extensions,
    we must prove the extensions degree is finite to conclude the extension is Galois.
    
    Suppose $|G| < [L : L^G]$. 
    Then there exists finite linearly independent set $S$ such that $|G| < |S|$. 
    Then since the $L^G$-subspace generated by $S$ is contained in $L^G(S)$, we have 
    \[ |G| < |S| = \dim_{L^G} \<S\>_{L^G} \leq [L^G(S) : L^G] \]
    Since all of $S$ is algebraic over $L^G$, 
    $L^G(S)$ is obtained by adjoining finitely many algebraic elements.
    So $L^G(S)$ is a finite extension, and thus a separable extension as well. 
    Hence there exists a primitive element $a$ where $L^G(S) = L^G(a)$. 
    But then \[
        |G| \leq |Orb(a)| = |\min(a,L^G)| = [L^G(a) : L^G] = [L^G(S) : L^G]
    \]
    which is a contradiction. 
    
    Thus, $[L : L^G] \leq |G|$. In particular, $L$ is a Galois $L^G$-extension.
    Furthermore, since $G \subseteq \gal{L^G}{L}$, \[
        |G| \leq |\gal{L^G}{L}| = [L : L^G] 
    \]
    So $[L : L^G] = |G|$. 
    We also have $|\gal{L^G}{L}| = |G|$. 
    Since both are finite sets and $G \subseteq \gal{L^G}{L}$,
    we have $G = \gal{L^G}{L}$. 
\end{proof}

\begin{dfn} 4 Equivalent Definitions of Galois Extension. 
    % Adapted from Milne's Galois Theory notes
    
    Let $\iota_L : K \to L$. 
    Let $\aut{K}{L}$ denote the $K$-extension automorphisms of $(L,\iota_L)$. 
    $\aut{K}{L}$ is a subgroup of 
    the group of ring automorphisms of $L$, $\aut{\Ring}{L}$. 
    For a subgroup $G$ of $\aut{\Ring}{L}$, 
    since $G \leq_\Grp \aut{\Set}{L}$, 
    we have a natural $G$ action on $L$ via $l \mapsto g(l)$ for $l \in L$. 
    Define \textbf{the subfield of $L$ fixed by $G$} as
    \[ L^G := \{l \in L \mid \forall g \in G, g(l)=l\} \]
    Then the following are equivalent: 
    \begin{enumerate}
        \item $(L,\iota_L)$ is a finite, normal and separable extension.
        \item $(L,\iota_L)$ is the splitting field 
            of some separable polynomial $f$ over $K$.
        \item $(L,\iota_L)$ is a finite extension such that $\bar{\iota_L}K = L^{\aut{K}{L}}$
        \item There exists a finite subgroup $G$ of $\aut{\Ring}{L}$ such that $\bar{\iota_L} K = L^G$.
    \end{enumerate}
    If any of the above is true,
    we call $(L,\iota_L)$ a \textbf{Galois $K$-extension}
    and $\aut{K}{L}$ is called the \textbf{Galois group of $(L,\iota_L)$} 
\end{dfn}
\begin{proof}
    ($1\imp 2$)
        
        There exists generators $a_0,\dots,a_{n-1}$ of $L$ as a $K$-extension.
        Let $f = \prod_{i\in n} \min(a_i,K)$. 
        Then $(L,\iota_L)$ algebraic, normal and separable implies 
        $L$ splits $f$ and $f$ is separable.
        Since the roots of $f$ contain the generators $a_0,\dots,a_{n-1}$,
        the roots generate $L$. Hence $L$ is the splitting field of $f$.
        
    ($2\imp 3$)
        
        $L$ generated by roots of $f$ implies 
        $L$ is generated by finitely many algebraic elements.
        So $L$ is a finite extension. 
        
        We leave the reader to check that $L^{\aut{K}{L}}$ indeed forms a subfield. 
        We wish to show $K = L^{\aut{K}{L}}$. 
        Clearly, $K \subseteq L^{\aut{K}{L}}$.
        We will show $[L : K] \leq [L : L^{\aut{K}{L}}]$ and hence
        by the tower law, $[K : L^{\aut{K}{L}}] = 1$, giving the desired result.
        
        We begin by noting that $\aut{K}{L} \subseteq \aut{L^{\aut{K}{L}}}{L}$.
        For all generators $a$ of $L$, 
        $\min(a,K)$ divides $f$ which is separable, so $\min(a,K)$ is separable. 
        By normality, for all generators $a$ of $L$, $L$ splits $\min(a,K)$.
        So by embedding via conjugates and all $\min(a,K)$ having distinct roots,
        we have $[L : K] = |\aut{K}{L}|$.
        Note that $L$ is also the splitting field of 
        $\bar{\iota_L} f$ considered as a polynomial over $L^{\aut{K}{L}}$
        and $\bar{\iota_L} f$ is still separable,
        so by the same argument we have 
        $[L : L^{\aut{K}{L}}] = |\aut{L^{\aut{K}{L}}}{L}|$. 
        Combining everything, we have \[
            [L : K] = |\aut{K}{L}| 
            \leq |\aut{L^{\aut{K}{L}}}{L}| = [L : L^{\aut{K}{L}}]
        \]
        as desired. 
        
    ($3\imp 4$)
        
        We claim that $\aut{K}{L}$ is finite and thus our desired subgroup $G$. 
        The key is to use embedding via conjugates again to bound 
        $|\aut{K}{L}|$ by $[L : K]$. 
        Since $[L : K]$ is finite, we have a finite set of generators of $L$.
        But we do not know whether $L$ splits 
        the minimal polynomials of these generators as required in the condition
        of embedding via conjugates. 
        So we take $(N,\iota_N)$, the normal closure of $(L,\iota_L)$ instead
        and apply embedding via conjugates to $(N,\iota_N\circ\iota_L)$,
        giving $|\mor{K}{L}{N}| \leq [L : K]$.
        But of course, $K$-extension automorphisms of $L$ give rise to 
        $K$-extension morphisms from $L$ to $N$ by composition with $\iota_N$.
        So $|\aut{K}{L}| \leq |\mor{K}{L}{N}| \leq [L : K]$ which is finite. 
        
    ($4\imp 1$)        
        This is the more difficult implication.
        Let $a \in L$. We claim that \[
        \bar\iota_L \min(a,K) = \prod_{b \in Orb(a)} (X - b)
        \]
        This shows all elements of $L$ have minimal polynomials 
        that $L$ splits and are separable. 
    
        We first show that the product is actually in $\bar{\iota_L}K[X] = L^G[X]$. 
        Let $g \in G$. Then \[
            \bar{g} \prod_{b \in Orb(a)} (X - b) 
            = \prod_{b \in Orb(a)} (X - g(b)) = \prod_{b' \in Orb(a)} (X - b')
        \]
        Hence the product is fixed by all $g \in G$. 
        Writing the product as $\sum_{k\in n} \la_k X^k$ where $\la_k \in L$,
        we have $\la_k \in L^G$, i.e. the product is indeed in $\bar{\iota_L}K[X]$. 
        This shows $a$ is algebraic, and hence $\min(a,K)$ exists.
        
        Since $\bar{\iota_L} K = L^G$, $\bar\iota_L \min(a,K) = \min(a,L^G)$.
        So we only need to prove the product is irreducible in $L^G[X]$.
        The product is clearly non-zero and non-unit. 
        Let $f, f_0$ be polynomials over $L^G$ such that \[
            f f_0 = \prod_{b \in Orb(a)} (X - b)
        \]
        Evaluating both sides at $a$, 
        we have $0 = ev_a (f)$ or $0 = ev_a(f_0)$. 
        WLOG $0 = ev_a (f)$. 
        Then since $f \in L^G[X]$ implies 
        for all $g \in G$, $\bar{g}(f) = f$, 
        we have \[
            0 = g(0) = g(ev_a f) = ev_{g(a)}(\bar{g} f) = ev_{g(a)} f
        \]
        Hence for all $b \in Orb(a)$, $b$ is a root of $f$, i.e.
        there exists a polynomial $h$ over $L$ such that \[
            h \prod_{b \in Orb(a)} (X - b) = f
        \]
        The by $L[X]$ is an integral domain, $1 = f_0 h$,
        which implies $\deg f_0 = 0$, giving $f_0$ is a unit in $L^G[X]$. 
        This proves irreducibility of the product in $L^G[X]$. 
        Hence it is equal to $\min(a,L^G) = \min(a,K)$. 
    
        Thus, once we prove $[L : K]$ is finite, it is both normal and separable. 
    
        Suppose $|G| < [L : K]$. 
        Then there exists finite linearly independent set $S$ such that $|G| < |S|$. 
        Then since the $K$-subspace generated by $S$ is contained in $K(S)$, 
        we have \[ |G| < |S| = \dim_{K} \<S\>_{K} \leq [K(S) : K] \]
        Since all of $S$ is algebraic over $K$, 
        $K(S)$ is obtained by adjoining finitely many algebraic elements.
        So $K(S)$ is a finite extension, and thus a separable extension as well. 
        Hence there exists a primitive element $a$ where $K(S) = K(a)$. 
        But then \[
            |G| \leq |Orb(a)| = |\min(a,K)| = [K(a) : K] = [K(S) : K]
        \]
        which is a contradiction. 
        Thus, $[L : K] \leq |G|$ which is finite. 
\end{proof}

%    - Thm - Galois Correspondance / Fund Thm Gal
%        -  Well Def
%        -  Inv.1 
%        -  Inv.2 
%        -  Order.1
%        -  Order.2
%        -  Degree
%        -  Equivariance
%        -  Normality
%        -  Quotient

\begin{thm}
Let $L/K$ be a finite galois extension. Then there is one to one correspondence between fields and galois groups.


\end{thm}
%    -  Cor - Fund Thm Alg
%        -  No non-trivial odd deg ext of R
%        -  No non-trivial even deg ext of C 


\end{document}